Numerical modelling of the Cordilleran ice sheet

### 2014-07-20 kappa.tex
Add abstract.

>>>>

\maketitle

\begin{abstract}

This doctoral thesis presents a study of the late glacial history of
north-western North America using numerical ice sheet modelling calibrated by
geological evidence.

The North Americam Cordillera is an immense maze of mountain ranges and
intermontane basins which once hosted an ice mass of the size of Greenland: the
Cordilleran ice sheet. Although less extensive than the former Laurentide and
Fennoscandian ice sheets, the Cordilleran ice sheet perhaps remains the least
understood Pleistocene ice sheet in terms of its extent, volume, and dynamics.
Because of the irregular topography of the ice sheet's subglacial landscape,
ice flowed along tortuous pathways, sometimes following the major topographical
troughs, and sometimes flowing across them.

This thesis presents simulations performed with PISM, a numerical ice sheet
model that approximates glacier flow by basal sliding and internal deformation.
These simulations show that the somewhat paradoxical geometry of the last
glacial maximum Cordilleran ice sheet can be largely explained by the present
distribution of temperature and precipitation in the region, provided an
accurate treatment of seasonality, orographic precipitation, and daily
temperature variability in the ice sheet model. Nevertheless, at the scale of
the last glacial cycle, this configuration appears ephemeral. In contrast, the
average stage of glaciation consists of isolated glaciers and ice caps covering
the major mountain ranges of the North American Cordillera.

The simulations presented in this thesis demonstrate the potential of combining
numerical ice sheet modelling to analysis of geological evidence in order to
improve our understanding former glaciations.

\end{abstract}

\newpage
